---
title: CV
name: Monica
surname: Thieu, PhD
position: "Postdoctoral scholar"
address: "Dept of Psychology, Emory University"
www: www.monicathieu.com
email: "mthieu@emory.edu"
twitter: monica_too_
github: monicathieu
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Marie is a Polish and naturalized-French physicist and chemist who conducts pioneering research on radioactivity."
headcolor: "#00ff00"
output: 
  vitae::markdowncv:
    theme: davewhipp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
require(tidyverse)
require(vitae)
```

## Currently

```{r}
tribble(
  ~ Position, ~ Year, ~ Institution, ~ Where, ~ Why,
  "Postdoctoral fellow", "August 2022 -", "Emory University", "Atlanta, GA, USA", c("Project: Subcortical pathways of looming and threat", "Supervisor: Dr. Philip Kragel", "Funding: [FIRST NIH IRACDA postdoctoral fellowship](https://med.emory.edu/education/postdoctoral-training/first/index.html)")
) %>% 
  # in the davewhipp theme, the where (4th) argument gets bolded. weirdly
  detailed_entries(Institution, Year, Where, Position, Why)
```

In my research, I use behavioral, modeling, and fMRI research methods to characterize how perceptual input contributes to affective experience.

In my teaching, I design and practice reflective, evidence-based, and student-centered training for psychological scientists, with a focus on technical training in statistics and scientific computing.

* Method competencies:
  * Functional neuroimaging: **Intermediate** in nipype, SPM, FSL; **proficient** in fmriprep
  * Behavioral tasks: **Advanced** in Gorilla, Qualtrics; **intermediate** in PsychoPy
  * Data management tools: **Intermediate** in REDCap
  * Coding: **Advanced** in R; **intermediate** in Python, Matlab, shell; **proficient** in JavaScript, Stan

## Education

```{r}
tribble(
  ~ Position, ~ Year, ~ Institution, ~ Where, ~ Why,
  "PhD in Psychology", "2022", "Columbia University", "New York, NY, USA", c("Dissertation: ['Exploring perception and categorization of social and affective stimuli'](https://academiccommons.columbia.edu/doi/10.7916/zjfd-fq65)", "Advisor: Dr. Kevin Ochsner", "Sep 2017 - May 2022"),
  "BA with honors in Psychology", "2015", "Stanford University", "Palo Alto, CA, USA", c("Honors thesis: 'Chronic media multitasking and distraction: Neural substrates and mnemonic consequences'", "Advisors: Dr. Melina  Uncapher, Dr. Anthony Wagner")
) %>% 
  detailed_entries(Institution, Year, Where, Position, Why)
```

## Research experience

```{r}
tribble(
  ~ Position, ~ Year, ~ Institution, ~ Where, ~ Why,
    "Postdoctoral fellow", "August 2022 -", "Emory University", "Atlanta, GA, USA", c("Project: Subcortical pathways of looming and threat", "Supervisor: Dr. Philip Kragel"),
  "Lab manager & research coordinator", "June 2015 - July 2017", "Stanford University", "Palo Alto, CA, USA", c("Project: Healthy aging and memory", "Supervisor: Dr. Anthony Wagner")
) %>% 
  detailed_entries(Institution, Year, Where, Position, Why)
```

## Awards and honors

### Research

```{r}
tribble(
  ~Year, ~Award, ~Desc,
  "2023", "[Diversity travel award](https://socialaffectiveneuro.org/wp-content/uploads/2023/05/SANS-Conference-2023_Program_FINAL.pdf) ($500)", "Social and Affective Neuroscience Society",
  "2022-26", "[FIRST NIH IRACDA postdoctoral fellowship](https://med.emory.edu/education/postdoctoral-training/first/index.html) (100% salary support for 4 years)", "Emory School of Medicine",
  "2019", "Honorable mention", "[NSF GRFP](https://nsfgrfp.org)",
  "June 2015", "Firestone Medal for Excellence in Undergraduate Research", "Stanford Vice Provost for Undergraduate Education",
  "Summer 2014", "Undergraduate Summer Research Award grant ($6,000)", "Stanford Bio-X",
  "Summer 2013", "Best Presentation", "Stanford Psych-Summer",	
  "Summer 2013", "Undergraduate research grant ($6,000)", "Stanford Psych-Summer",
  
) %>% 
  brief_entries(
    Award,
    Year, 
    Desc
  )
```

### Teaching

```{r}
tribble(
  ~Year, ~Award, ~Desc,
  "Spring 2022", "[Teaching Development Program Advanced Track certificate](https://tdp.ctl.columbia.edu/about/overview/)", "Columbia Center for Teaching and Learning",
  "Spring 2022", "Presidential Graduate Teaching Award finalist", "Columbia Office of the Provost",
  "Spring 2021", "[Grant to Enhance Remote Learning and Teaching ($1,500)](https://scienceandsociety.columbia.edu/content/remote-learning-grants-overview)", "Columbia Center for the History and Ethics of Public Health",
  "Spring 2020", "[Curricular Innovation Fellow](https://rcfoundations.research.columbia.edu/content/foundations-curriculum-innovation-grant-program)", "Columbia Libraries",
  "June 2015", "Zimbardo Teaching Prize", "Stanford Psych One",		
) %>% 
  brief_entries(
    Award,
    Year, 
    Desc
  )
```

### Other

```{r}
tribble(
  ~Year, ~Award, ~Desc,
  "Mar 2024", "Quarterfinalist", "Jeopardy! Invitational Tournament",
  "Aug 2023 - present", "Trivia expert", "Master Minds (Game Show Network)",
  "Mar 2019", "Second place runner-up (team)", "Jeopardy! All-Star Games",
  "May 2016", "Contestant", "500 Questions (ABC)",
  "Feb 2013", "Quarterfinalist", "Jeopardy! Tournament of Champions",
  "Feb 2012", "First place winner", "Jeopardy! College Championship",
  "Winter 2013", "Excellence Award", "Stanford Introductory Seminars",
) %>% 
  brief_entries(
    Award,
    Year, 
    Desc
  )
```

## Publications

```{r}
bibliography_entries("publications.bib") # %>% 
# appears not to work when there are null date papers...?
  # arrange(desc(issued))
```

## Presentations

### Talks and workshops

```{r}
tribble(
  ~Year, ~Talk, ~Conference,
  "March 2024", "Advanced Data Visualization with ggplot2", "Emory Graduate Students in Psychology and Neuroscience workshop",
  "April 2023", "Shallow neural networks for collision detection predict arousal-related differences in subjective experience", "Social and Affective Neuroscience Society Annual Meeting",
    "Jan 2023", "Remote brain bingo: Running group games on Zoom", "TIP Coast-To-Coast Conference",
  "June 2022", "Remote brain bingo: Running group games on Zoom", "Psych One Conference",
  "July 2020", "How to Prepare and Teach an R Lesson", "[NYC R Conference](https://rstats.ai/nyr/#speakers)",
  "Nov 2020", "Exploring the NYPD Civilian Complaint Database", "Columbia Foundations for Research Computing workshop",
  "Apr 2020", "Wrangling Multilevel Data in the Tidyverse", "Columbia Foundations for Research Computing workshop",
) %>% 
  brief_entries(
    Talk,
    Year, 
    Conference
  )
```

### Conference posters

```{r}
bibliography_entries("posters.bib") %>% 
  arrange(desc(issued))
```

## Teaching experience

### Course instructor

```{r}
tribble(
  ~ Position, ~ Year, ~ Course, ~ Where,
  "Instructor of record", "Sp2024", "PSY 417: Statistics II in Psychology", "Spelman College Dept of Psychology",
  "Instructor", "Su2023, Su2024", "Data Science Lab Module", "[Spelman College WiSTEM](https://www.spelman.edu/academics/research-programs/wistem)",
  "Instructor of record", "Su2022", "PSYC 1021: Science of Psychology: Explorations & Applications", "Columbia Dept of Psychology"
) %>% 
  # flipped Where and Institution to get the institution to be bolded
  detailed_entries(Course, Year, Where, Position)
```

### Workshop instructor

```{r}
tribble(
  ~ Position, ~ Year, ~ Course, ~ Where,
  "[Carpentries](https://carpentries.org/)-certified instructor", "Fa2019-Sp2022", "Introductory coding bootcamps", "[Columbia Foundations for Research Computing](https://rcfoundations.research.columbia.edu/)",
  "Curriculum developer & instructor", "Sp2018-Fa2021", "[Intro to Programming for Social Scientists](https://cu-psych-computing.github.io/cu-psych-comp-tutorial/workshop/)", "Columbia Dept of Psychology"
) %>% 
  # flipped Where and Institution to get the institution to be bolded
  detailed_entries(Course, Year, Where, Position)
```

### Guest lecturer, section leader, & teaching assistant

```{r}
tribble(
  ~ Position, ~ Year, ~ Course, ~ Where,
  "Assistant curriculum developer & TA", "Sp2022", "PSYC 2821: Cognition in Context: Mental Health & Well-being", "Columbia Dept of Psychology",
  "Assistant curriculum developer, TA, & guest lecturer", "Sp2021", "PSYC 1021: Science of Psychology: Explorations & Applications", "Columbia Dept of Psychology",
  "Guest lecturer", "Fa2019", "[Intro to Psychology](https://www.outlier.org/products/intro-to-psychology)", "Outlier.org",
  "TA & guest lecturer", "Sp2019-Fa2021", "B8144: Intro to Programming in R", "Columbia Business School",
  "TA & section leader", "Fa2018, Fa2019", "PSYC 1490: Research Methods", "Columbia Dept of Psychology",
  "TA & guest lecturer", "Fa2017", "PSYC 1001: Science of Psychology", "Columbia Dept of Psychology",
  "Teaching fellow", "Fa2014-Sp2015", "Psych One", "Stanford Dept of Psychology"
) %>% 
  # flipped Where and Institution to get the institution to be bolded
  detailed_entries(Course, Year, Where, Position)
```

## Popular media

```{r}
tribble(
  ~Year, ~Title, ~Publication,
  "May 2024", "[Radio/podcast guest (trivia, science)](https://www.npr.org/transcripts/1196979444)", "Life Kit by NPR",
  "2021-2022", "Host of [The Psychology PhD podcast](https://psychology.columbia.edu/podcast)", "Columbia Dept of Psychology",
  "Nov 2021", "[Video guest (science)](https://youtu.be/VTvJOb2hJ6k)", "Wired 5 Levels: Memory",
  "Sept 2021", "[Interview subject (teaching)](https://explaineverything.com/psychology-whiteboard-video-lessons-with-monica-thieu/)", "Explain Everything",
  "July 2020", "[Podcast guest (science)](https://www.marketplace.org/shows/this-is-uncomfortable-reema-khrais/money-makes-me-sick/)", "This is Uncomfortable by Marketplace",
  "July 2019", "[Podcast guest (trivia, statistics)](https://www.npr.org/transcripts/722198188)", "Planet Money by NPR"
) %>% 
  brief_entries(
    Title,
    Year, 
    Publication
  )
```

## Mentoring

```{r}
tribble(
  ~ Name, ~ Year, ~ Where, ~ Position, ~ Awards,
  "Lauren Wilkins", "2020-22", "Columbia University", "Undergraduate research assistant & honors thesis student", "Class of 1939 Summer Research Fellowship, summer 2020"
) %>% 
  detailed_entries(Position, Year, Where, Name, Awards)
```

## Service

```{r}
tribble(
  ~ Position, ~ Year, ~ Org, ~ Where, ~ Details,
  "Treasurer", "2023 - present", "Emory Postdoctoral Association", "Emory University", "",
  "Peer reviewer", " ", " ", "ad hoc", "Cerebral Cortex; Neuropsychologia; Social, Cognitive, & Affective Neuroscience"
) %>% 
  detailed_entries(Org, Year, Where, Position, Details)
```


## Workshops & summer schools attended

```{r}
tribble(
  ~Year, ~Workshop, ~Desc,
  "2022", "[Neurohackademy](https://neurohackademy.org)", "University of Washington",
  "2019", "[Kavli Summer Institute in Cognitive Neuroscience](https://sicn.cmb.ucdavis.edu/)", "University of California, Santa Barbara",
  "2018", "São Paulo School of Advanced Sciences on Social & Affective Neuroscience", "Mackenzie Presbyterian University"
) %>% 
  brief_entries(
    Workshop,
    Year, 
    Desc
  )
```
